Theoretical aspects in gravity modelling

### 2018-05-10 manuscript/letter.tex
Review of the introduction with slight changes
I reworded some parts of the introduction to make it read more smoothly.
Removed some repetition and added more idioms.

\author[Oliveira Jr et al.]
{Vanderlei C. Oliveira Jr$^1$\thanks{Email: vanderlei@on.br}, 
Kristoffer A. T. Hallam$^{1}$, Val\'{e}ria C. F. Barbosa$^1$
and Leonardo Uieda$^2$ \\
$^1$ Department of Geophysics, Observat\'{o}rio Nacional, Rio de Janeiro, Brazil \\
$^2$ Department of Geology and Geophysics, University of Hawaii, Manoa, USA
}

@@

\begin{summary}
 The gravity anomaly has long been used by geophysicists for 
 representing the gravitational effect produced by gravity
 sources.This quantity, however, can be used in this way as a good
 approximation. Rigorously, the gravitational effect produced by the sources
 is represented by the gravity disturbance, which is a well-known 
 quantity in geodesy.
 In this paper, we discuss the fundamental concepts supporting 
 the fact that gravity disturbance is the more appropriated quantity
 to be used for estimating density variations related to geological
 bodies in subsurface. We show that the maximum absolute difference 
 between the gravity disturbance and the Free-air anomaly in Brazil 
 reaches a maximum absolute value of $\approx 9.3$ mGal. We also stress 
 that almost all authors in the literature have implicitly used the 
 gravity anomaly as an approximation of the gravity disturbance, 
 which can be represented by a harmonic function and, consequently, 
 satisfies the premise behind most of 
 the interpretation techniques used in geophysics (e.g., 
 upward/downward continuation, data processing with equivalent layer,
 conversions between gravity and magnetic data, computation of vertical 
 derivatives via Fourier and Hilbert transforms).
 Although the gravity anomaly can be considered a good approximation of the 
 gravity disturbance for most practical applications, 
 it is important to bear in mind not only the terminology 
 changes, but also the conceptual assumptions.

@@

The resultant of gravitational force and centrifugal force acting 
on a body at rest on the Earth's surface is called 
\textit{gravity vector} and its intensity is called 
\textit{gravity} \citep{heiskanen-moritz1967,
hofmann-wellenhof-moritz2005}.
In the case of gravimetry on moving platforms (e.g., airplanes,
helicopters, marine vessels), there are additional
non-gravitational accelerations due to the vehicle motion, 
such as Coriolis acceleration and high-frequency vibrations 
\citep{glennie-etal2000,nabighian-etal2005-grav,baumann-etal2012}.
Geophysicists use gravity for estimating the Earth's 
internal density distribution whereas geodesists use
gravity to estimate the Geoid \citep{li2001}.
Hence, geophysicists are usually interested 
in the gravitational component of the observed gravity, 
which is produced by the Earth's internal 
density distribution.

The first step of the procedure for isolating this 
gravitational component consists in removing the non-gravitational 
effects due to the vehicle motion and also the time variations 
such as Earth tides, instrumental drift and barometric 
pressure changes, for example.
If these effects are properly removed, the resultant 
gravity data can be considered as the sum of a 
centrifugal component due to the Earth's rotation and
a gravitational component produced by the whole Earth's
internal density distribution.
The isolation of this particular gravitational component 
and its subsequent use for estimating density 
distributions related to geological structures in subsurface 
are the main goals in applied geophysics \citep{blakely1996}.
The computation of the gravitational effect produced by
the geological structures is commonly known in geophysics as 
\textit{gravity modelling}.
This term is different from \textit{gravity field modelling}, which
is commonly used in geodesy to denote the characterization of the 
gravity field in a local, regional or global scale.

Based on well-established concepts of the literature,
we present a discussion aiming at bringing some
light to the following question: in geophysical applications,
should we use the gravity disturbance or gravity anomaly?
It seems that this theoretical issue has been 
debated within the scientific community from a 
more geodetic than geophysical point of view
\citep{lafehr1991,chapin1996,li2001,fairhead2003,
hackney-featherstone2003,hinze2005}.
Our reasoning suggests that the gravity 
disturbance is more appropriated than the gravity anomaly 
for geophysical purposes.

@@

\item Almost all forward modelling techniques compute
the vertical component of the gravitational attraction 
exerted by the gravity sources at the observation points.
Notice that the gravity anomaly requires the computation of gravity
on the Geoid, which is generally within the topographic masses.
Hence, almost all these forward modelling techniques in fact
compute the gravity disturbance.

>>>>

\author[Oliveira Jr et al.]
{Vanderlei C. Oliveira Jr$^1$\thanks{Email: vanderlei@on.br},
Kristoffer A. T. Hallam$^{1}$, Val\'{e}ria C. F. Barbosa$^1$
and Leonardo Uieda$^2$ \\
$^1$ Department of Geophysics, Observat\'{o}rio Nacional, Rio de Janeiro, Brazil \\
$^2$ Department of Geology and Geophysics, SOEST, University of Hawai'i at M\={a}noa, Honolulu, USA
}

@@

\begin{summary}
 The gravity anomaly has long been used by geophysicists for
 representing the gravitational effect produced by gravity
 sources. This quantity, however, can be used in this way as a good
 approximation. Rigorously, the gravitational effect produced by the sources
 is represented by the gravity disturbance, which is a well-known
 quantity in geodesy.
 In this paper, we discuss the fundamental concepts supporting
 the fact that gravity disturbance is the more appropriated quantity
 to be used for estimating density variations related to geological
 bodies in subsurface. We show that the maximum absolute difference
 between the gravity disturbance and the Free-air anomaly in Brazil
 reaches a maximum absolute value of $\approx 9.3$ mGal. We also stress
 that almost all authors in the literature have implicitly used the
 gravity anomaly as an approximation of the gravity disturbance,
 which can be represented by a harmonic function and, consequently,
 satisfies the premise behind most of
 the interpretation techniques used in geophysics (e.g.,
 upward/downward continuation, data processing with equivalent layer,
 conversions between gravity and magnetic data, computation of vertical
 derivatives via Fourier and Hilbert transforms).
 Although the gravity anomaly can be considered a good approximation of the
 gravity disturbance for most practical applications,
 it is important to bear in mind not only the terminology
 changes, but also the conceptual assumptions.

@@

The \textit{gravity vector} is the sum of the gravitational and centrifugal
accelerations felt by a body at rest on the Earth's surface.
Its intensity is what is geoscientists call \textit{gravity}
\citep{heiskanen-moritz1967, hofmann-wellenhof-moritz2005}.
When measured on moving platforms (e.g., airplanes,
helicopters, marine vessels), there are additional
non-gravitational accelerations due to the vehicle motion,
such as Coriolis acceleration and high-frequency vibrations
\citep{glennie-etal2000,nabighian-etal2005-grav,baumann-etal2012}.
Geodesists use gravity measurements to estimate the Geoid \citep{li2001}
whereas geophysicists use gravity to estimate the Earth's
internal density distribution.
Hence, geophysicists are usually only interested
in the gravitational component of the observed gravity
because it is the part that reflects the Earth's internal density distribution.

The first step toward isolating the gravitational component
is to remove the effects of vehicle motion, Earth tides, instrumental drift,
and barometric pressure changes, among others.
If these effects are properly removed, the resultant
observations can be considered to be solely the sum of the
centrifugal acceleration due to the Earth's rotation and
the gravitational acceleration produced by
the internal density distribution of the whole Earth.
The isolation of this particular gravitational component
and its subsequent use for estimating density
distributions related to geological structures in subsurface
are the main goals of applied gravimetry \citep{blakely1996}.
The computation of the gravitational effect produced by
the geological structures is commonly known in geophysics as
\textit{gravity modelling}.
This term is different from \textit{gravity field modelling}, which
is commonly used in geodesy to denote the mathematical characterization of the
gravity field in local, regional, or global scales (e.g., spherical harmonic
analysis).

We present a discussion with the aim of shedding some light on the question:
\textit{should we use the gravity disturbance or gravity anomaly in geophysical
applications}?
The concepts discussed here are well-established in the literature.
However, the debate around this theoretical issue has been
carried about from a more geodetic than a geophysical point of view
\citep{lafehr1991,chapin1996,li2001,fairhead2003,
hackney-featherstone2003,hinze2005}.
Our reasoning suggests that the gravity disturbance is more appropriated for
geophysical purposes than the gravity anomaly.

@@

\item Almost all forward modelling techniques compute
the vertical component of the gravitational attraction
exerted by the gravity sources at the observation points.
Notice that the gravity anomaly requires the computation of gravity
on the Geoid, which is generally within the topographic masses.
Hence, almost all these techniques in fact compute the gravity
disturbance.
